---
title: "Visium Spatial Tx - Mouse Hippocampus (hip142079) - Manuscript figures"
author: '[Kevin Lebrigand](mailto:lebrigand@ipmc.cnrs.fr) CNRS Senior Scientist - Computational Biology'
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  html_notebook:
    code_folding: none
    theme: journal
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

<style type="text/css">

body, td {
   font-size: 15px;
}
code.r{
  font-size: 15px;
}
pre {
  font-size: 15px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  cache.lazy = FALSE,
  tidy = TRUE
)

```

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

setwd("/data/10x_data/10x_visium/")
source("00.import.R")

mob164315 <- readRDS("./output/mob164315.nanopore.rds")
hip142078 <- readRDS("./output/hip142078.nanopore.rds")
hip142079 <- readRDS("./output/hip142079.nanopore.rds")
clusters <- c("Midbrain", "Hippocampus", "Isocortex-1", "Isocortex-2", "Olfactory area", "Fiber tracts", "Isocortex-3", "Hammons horn", "Thalamus", "Hypothalamus", "Ca3", "Dg")

```

# multi-isoform genes ratio (not retained)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

#hip142079 <- subset(hip142079, subset = nFeature_ISO>1)
#mm <- numeric()
#for(i in 1:2500){
#  t <- sapply(strsplit(rownames(hip142079@assays$ISO@counts[hip142079@assays$ISO@counts[,i]>0,]), "\\.\\."), `[`, 1)
#  multi <- length(t) - length(unique(t))
#  ratio <- 100*multi/length(t)
#  print(paste(i,length(t),multi,ratio))
#  mm <- c(mm,ratio)
#}

#names(mm) <- colnames(hip142079@assays$ISO@counts)
#hip142079[['multi']] <- mm
#hip142079[['multinorm']] <- hip142079[['multi']]/hip142079[['nCount_Spatial']]
#SpatialFeaturePlot(hip142079, features=c("multi","multinorm","nCount_Spatial"), max.cutoff=c(10,0.0015,30000))
#ggplot(data=hip142079@meta.data, aes(x=nCount_Spatial, y=multi)) +
#    geom_point(alpha=0.5, shape=21, color="black") +
#    theme_minimal() +
#    theme(legend.position="bottom") +
#    ylab("multi") +
#    xlab("nCount") +
#    theme(legend.position = "none")

```

# Supplementary figure Sept5 (not retained)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

tiff("figures/dotplot.sept5.spatial.tif", units="in", width=6, height=6, res=300)
DefaultAssay(hip142079) <- "Spatial"
SpatialFeaturePlot(hip142079, features=c("Sept5","Gp1bb"),max.cutoff = 3)
dev.off()
tiff("figures/dotplot.sept5.isog.tif", units="in", width=6, height=6, res=300)
DefaultAssay(hip142079) <- "ISOG"
SpatialFeaturePlot(hip142079, features=c("Sept5","Gp1bb"),max.cutoff = 3)
dev.off()

```

# Supplementary statistics (not retained)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

#library(fmsb)
# Create data: note in High school for several students
#set.seed(99)
#data <- as.data.frame(matrix( sample( 0:20 , 15 , replace=F) , ncol=5))
#colnames(data) <- c("math" , "english" , "biology" , "music" , "R-coding" )
#rownames(data) <- paste("mister" , letters[1:3] , sep="-")
 
# To use the fmsb package, I have to add 2 lines to the dataframe: the max and min of each variable to show on the plot!
#data <- rbind(rep(20,5) , rep(0,5) , data)
 
# plot with default options:
#radarchart(data)
#x <- read.table("stats.txt", sep="\t")

#rownames(x) <- c("1","2","hip142079.1","hip142079.2","hip142079.3","hip142079.4","hip142079.5","hip142079.6","hip142079.7","hip142078.1","hip142078.2","mob164315.1","mob164315.2")
#colnames(x) <- c("PolyA found","cellBC found","UMI found","UMI genes found")

#radarchart(x)


#colors_border=c( "#2af598", "#22e4ac",  "#1bd7bb", "#14c9cb" , "#0fbed8", "#08b3e5", "#0F496E", "#BC438A", "#F3D8E8" , "#cb5323", "#591702" )
#colors_in=c( "#2af598", "#22e4ac",  "#1bd7bb", "#14c9cb" , "#0fbed8", "#08b3e5", "#0F496E", "#BC438A", "#F3D8E8" , "#cb5323", "#591702" )

#radarchart( x  , axistype=1 , 
    #custom polygon
#    pcol=colors_border , plwd=4 , plty=1,
    #custom the grid
#    cglcol="grey", cglty=1, axislabcol="grey", caxislabels=seq(0,100,25), cglwd=0.8,
    #custom labels
#    vlcex=0.8 
#    )

```

# Supplementary fig.3

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

DefaultAssay(object = hip142079) <- "ISO"
general <- c('Snap25','Cnih2','Gnas','Bin1','Caly','Dtnbp1','Aldoa')
lst <- c()
for (i in (1:length(general))){
  x <- rownames(hip142079@assays$ISO@counts)[grep(paste("^",general[i],"\\.\\.",sep=""), rownames(hip142079@assays$ISO@counts))]
  lst <- c(lst,x)
}
#iso.switching = read.delim("./output/hip142079.isoswitch.top.csv", sep=",", stringsAsFactors = F)
#iso.switching$key <- paste(iso.switching$geneId,iso.switching$transcriptId, sep="..")
#lst <- lst[lst %in% iso.switching$key == TRUE]

tiff("figures/dotplot.hip.iso.tif", units="in", width=16, height=8, res=300)
DefaultAssay(object = hip142079) <- "ISO"
DotPlot(object = hip142079, features = lst, scale=TRUE, scale.max = 100, scale.min = 0, col.min = -1, col.max = 2) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
dev.off()

tiff("figures/dotplot.hip.gene.tif", units="in", width=24, height=6, res=300)
DefaultAssay(object = hip142079) <- "SCT"
print(DotPlot(object = hip142079, features = general, scale=TRUE, scale.max = 100, scale.min = 0, col.min = -1, col.max = 2, cols = c("lightgrey", "red")) + theme(axis.text.x = element_text(angle = 90, hjust = 1))) # 
dev.off()

DefaultAssay(object = hip142079) <- "ISOG"
tiff("figures/dotplot.hip.isog.tif", units="in", width=16, height=6, res=300)
print(DotPlot(object = hip142079, features = general, scale.max = 100, scale.min = 0, col.min = -1, col.max = 2, cols = c("lightgrey", "purple")) + theme(axis.text.x = element_text(angle = 90, hjust = 1))) # 
dev.off()

# mob
mob164315 <- readRDS("./output/mob164315.nanopore.rds")
DefaultAssay(object = mob164315) <- "ISO"
general <- c('Myl6','Plp1')
lst <- c()
for (i in (1:length(general))){
  x <- rownames(mob164315@assays$ISO@counts)[grep(paste("^",general[i],"\\.\\.",sep=""), rownames(mob164315@assays$ISO@counts))]
  lst <- c(lst,x)
}
iso.switching = read.delim("./output/mob164315.isoswitch.top.csv", sep=",", stringsAsFactors = F)
iso.switching$key <- paste(iso.switching$geneId,iso.switching$transcriptId, sep="..")
lst <- lst[lst %in% iso.switching$key == TRUE]

tiff("figures/dotplot.mob.iso.tif", units="in", width=6, height=6, res=300)
DefaultAssay(object = mob164315) <- "ISO"
DotPlot(object = mob164315, features = lst, scale=TRUE, scale.max = 100, scale.min = 0) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
dev.off()

tiff("figures/dotplot.mob.gene.tif", units="in", width=6, height=4, res=300)
DefaultAssay(object = mob164315) <- "SCT"
print(DotPlot(object = mob164315, features = general, scale.max = 100, scale.min = 0, cols = c("lightgrey", "red")) + theme(axis.text.x = element_text(angle = 90, hjust = 1))) # 
dev.off()

DefaultAssay(object = mob164315) <- "ISOG"
tiff("figures/dotplot.mob.isog.tif", units="in", width=6, height=4, res=300)
print(DotPlot(object = mob164315, features = general, scale.max = 100, scale.min = 0, col.min = -1, col.max = 2,cols = c("lightgrey", "purple")) + theme(axis.text.x = element_text(angle = 90, hjust = 1))) # 
dev.off()

```

# Main fig.2b

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

#pdf("figures/fig2a.pdf", width=36, height=8, useDingbats=FALSE)
#SpatialFeaturePlot(hip142079, features = c("AtoI_A","AtoI_G","AtoI_ratio"), pt.size.factor=2, alpha = c(1, 1), max.cutoff = c(300,80,0.6))
#dev.off()

tiff("figures/fig2d.tif", units="in", width=20, height=8, res=300)
SpatialFeaturePlot(hip142079, features = c("AtoI_A","AtoI_G","AtoI_ratio"), pt.size.factor=1.6, alpha = c(1, 1), max.cutoff = c(300,80,0.4))
dev.off()

```

# Main fig.2c

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

col=ggplotColours(n = length(unique(hip142079@active.ident)))

ggplot(data=hip142079@meta.data, aes(x=ClusterName,y=AtoI_ratio, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_y_continuous(limits=c(0,0.4), breaks=seq(0,0.4,0.1), expand = c(0, 0))

tiff("figures/fig.2c.tif", units="in", width=8, height=6, res=300)
ggplot(data=hip142079@meta.data, aes(x=ClusterName,y=AtoI_ratio, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_y_continuous(limits=c(0,0.4), breaks=seq(0,0.4,0.1), expand = c(0, 0))
dev.off()

ggplot(data=hip142079@meta.data, aes(x=ClusterName,y=AtoI_detected, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_y_continuous(limits=c(0,50), breaks=seq(0,50,10), expand = c(0, 0))

```

# Main fig.2c-2

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

avg <- AverageExpression(object = hip142079, return.seurat = TRUE, assay="SCT")
pheatmap(avg@assays$SCT@scale.data[c("Adar","Adarb1","Adarb2"),], cluster_rows=FALSE, cluster_cols=FALSE, fontsize = 8, clustering_method = "complete", clustering_distance_rows = "manhattan")

#pdf("figures/fig2c.pdf", width=12, height=6, useDingbats=FALSE)
#pheatmap(avg@assays$SCT@scale.data[c("Adar","Adarb1","Adarb2"),], cluster_rows=FALSE, cluster_cols=FALSE, fontsize = 8, clustering_method = "complete", clustering_distance_rows = "manhattan")
#dev.off()

tiff("figures/fig2c-2.tif", units="in", width=8, height=6, res=300)
pheatmap(avg@assays$SCT@scale.data[c("Adar","Adarb1","Adarb2"),], cluster_rows=FALSE, cluster_cols=FALSE, fontsize = 8, clustering_method = "complete", clustering_distance_rows = "manhattan")
dev.off()

```

# Main fig.2x (not retained)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

hip142079[['SCTAdar']] <- hip142079@assays$SCT@data["Adar",]
hip142079[['SCTAdarb1']] <- hip142079@assays$SCT@data["Adarb1",]
hip142079[['SCTAdarb2']] <- hip142079@assays$SCT@data["Adarb2",]

p1 <- ggplot(data=hip142079@meta.data, aes(x=SCTAdar,y=AtoI_ratio, group=SCTAdar))+
  geom_boxplot(outlier.colour=NA, fill='#466cb9', color="black", lwd=0.1)+
  scale_y_continuous(limits=c(0,0.4))

p2 <- ggplot(data=hip142079@meta.data, aes(x=SCTAdarb1,y=AtoI_ratio, group=SCTAdarb1))+
  geom_boxplot(outlier.colour=NA, fill='#466cb9', color="black", lwd=0.1)+
  scale_y_continuous(limits=c(0,0.4))

p3 <- ggplot(data=hip142079@meta.data, aes(x=SCTAdarb2,y=AtoI_ratio, group=SCTAdarb2))+
  geom_boxplot(outlier.colour=NA, fill='#466cb9', color="black", lwd=0.1) +
  scale_y_continuous(limits=c(0,0.4))

plot_grid(p1,p2,p3,ncol=3)

hip142079[['UMIAdar']] <- hip142079@assays$Spatial@counts["Adar",]
hip142079[['UMIAdarb1']] <- hip142079@assays$Spatial@counts["Adarb1",]
hip142079[['UMIAdarb2']] <- hip142079@assays$Spatial@counts["Adarb2",]

p1 <- ggplot(data=hip142079@meta.data, aes(x=UMIAdar,y=AtoI_ratio, group=UMIAdar))+
  geom_boxplot(outlier.colour=NA, fill='#466cb9', color="black", lwd=0.1)+
  scale_y_continuous(limits=c(0,0.35))

p2 <- ggplot(data=hip142079@meta.data, aes(x=UMIAdarb1,y=AtoI_ratio, group=UMIAdarb1))+
  geom_boxplot(outlier.colour=NA, fill='#466cb9', color="black", lwd=0.1)+
  scale_y_continuous(limits=c(0,0.35))

p3 <- ggplot(data=hip142079@meta.data, aes(x=UMIAdarb2,y=AtoI_ratio, group=UMIAdarb2))+
  geom_boxplot(outlier.colour=NA, fill='#466cb9', color="black", lwd=0.1) +
  scale_y_continuous(limits=c(0,0.35))

plot_grid(p1,p2,p3,ncol=3)

#tiff("figures/fig2d.tif", units="in", width=12, height=6, res=300)
#plot_grid(p1,p2,p3,ncol=3)
#dev.off()

```

# Main fig.2x - per region (not retained)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

  p1 <- hip142079@meta.data %>% 
    ggplot(aes(x=factor(UMIAdar),y=AtoI_ratio, fill=ClusterName)) +
    geom_boxplot() +
    scale_y_continuous(limits=c(0,0.4)) +
    geom_jitter(width=0.1,alpha=0.2) +
    xlab("Adar")+ 
    facet_wrap(~ClusterName,ncol = 4) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  p2 <- hip142079@meta.data %>% 
    ggplot(aes(x=factor(UMIAdarb1),y=AtoI_ratio, fill=ClusterName)) +
    geom_boxplot() +
    scale_y_continuous(limits=c(0,0.4)) +
    geom_jitter(width=0.1,alpha=0.2) +
    xlab("Adarb1")+ 
    facet_wrap(~ClusterName,ncol = 4) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  p3 <- hip142079@meta.data %>% 
    ggplot(aes(x=factor(UMIAdarb2),y=AtoI_ratio, fill=ClusterName)) +
    geom_boxplot() +
    scale_y_continuous(limits=c(0,0.4)) +
    geom_jitter(width=0.1,alpha=0.2) +
    xlab("Adarb2")+ 
    facet_wrap(~ClusterName,ncol = 4) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  #tiff("figures/fig2d.tif", units="in", width=20, height=6, res=300)
  #plot_grid(p1,p2,p3,ncol=3)
  #dev.off()
  
```

# Supplementary single cell deconvolution per region (hip79, not retained)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

avg <- AverageExpression(object = hip142079, return.seurat = TRUE, assay="predClusterName")
avg <- ScaleData(avg, assay="predClusterName")
kable(head(sort(avg@assays$predClusterName@scale.data[,"Midbrain"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Hippocampus"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Isocortex-1"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Isocortex-2"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Olfactory area"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Fiber tracts"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Isocortex-3"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Fiber tracts"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Hammons horn"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Fiber tracts"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Thalamus"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Hypothalamus"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Ca3"], decreasing=TRUE),5))
kable(head(sort(avg@assays$predClusterName@scale.data[,"Dg"], decreasing=TRUE),5))

zeisel <- readRDS("./output/zeisel.rds")
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[1]], decreasing=TRUE),5))))
p1 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[1]))  + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[2]], decreasing=TRUE),5))))
p2 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[2])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[3]], decreasing=TRUE),5))))
p3 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[3])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[4]], decreasing=TRUE),5))))
p4 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[4])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[5]], decreasing=TRUE),5))))
p5 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[5])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[6]], decreasing=TRUE),5))))
p6 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[6])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[7]], decreasing=TRUE),5))))
p7 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[7])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[8]], decreasing=TRUE),5))))
p8 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[8])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[9]], decreasing=TRUE),6))))
p9 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[9])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[10]], decreasing=TRUE),5))))
p10 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[10])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[11]], decreasing=TRUE),5))))
p11 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[11])) + NoLegend()
sub <- subset(zeisel, subset = ClusterName %in% c(names(head(sort(avg@assays$predClusterName@scale.data[,clusters[12]], decreasing=TRUE),5))))
p12 <- DotPlot(sub, features=c("Adar","Adarb1","Adarb2"), group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[12])) + NoLegend()

plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,ncol=4)

```

# Main fig.1e (hip, Snap25 isoforms)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

x <- unique(rownames(hip142079@assays$ISO@data)[grep(paste("^Snap25..",sep=""), rownames(hip142079@assays$ISO@data))])
DefaultAssay(hip142079) <- "ISO"
SpatialFeaturePlot(hip142079, features = x, pt.size.factor=1.6, alpha = c(1, 1))

```

# Main fig.1b (mob, Plp1 isoforms)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

x <- unique(rownames(mob164315@assays$ISO@data)[grep(paste("^Plp1..",sep=""), rownames(mob164315@assays$ISO@data))])
DefaultAssay(mob164315) <- "ISO"
SpatialFeaturePlot(mob164315, features = x, pt.size.factor=1.6, alpha = c(1, 1))

```

# Supplementary fig.8 single cell deconvolution per region (mob, Plp1 story)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

mob164315 <- readRDS("./output/mob164315.nanopore.rds")
clusters <- c("Granule Cell Layer (GCL+RMS)","Mitral Cell Layer (MCL)","Outer plexiform Layer (EPL)","Glomerular Layer (GL)","Olfactory Nerve Layer (ONL)")
col=ggplotColours(n = length(unique(mob164315@active.ident)))

x <- unique(rownames(mob164315@assays$ISO@data)[grep(paste("^Plp1..",sep=""), rownames(mob164315@assays$ISO@data))])
DefaultAssay(mob164315) <- "ISO"
SpatialFeaturePlot(mob164315, features = x, pt.size.factor=1.6, alpha = c(1, 1))

#DefaultAssay(mob164315) <- "predictions"
#p1 = DotPlot(mob164315, features = sort(rownames(mob164315@assays$predictions@data))) + RotatedAxis()
#p1 + theme(axis.text.x = element_text(size = 6)) + xlab("Cluster")
#p1

avg <- AverageExpression(object = mob164315, return.seurat = TRUE, assay="predictions")
avg <- ScaleData(avg, assay="predictions")

kable(head(sort(avg@assays$predictions@scale.data[,"Granule Cell Layer (GCL+RMS)"], decreasing=TRUE),10))
kable(head(sort(avg@assays$predictions@scale.data[,"Olfactory Nerve Layer (ONL)"], decreasing=TRUE),10))

tepe <- readRDS("./output/tepe.rds")
p1 <- DimPlot(tepe, group.by="ClusterName", label=TRUE)
p2 <- FeaturePlot(tepe, "Plp1")

GCL.clust <- c("N4","N10","N8","N9","N7","N11","N1","OPC")
ONL.clust <- c("Mural1","Mes2","OEC5","OEC2","OEC4","Mes1","OEC1","OEC3")

sub <- subset(tepe, subset = ClusterName %in% c(names(head(sort(avg@assays$predictions@scale.data[,clusters[1]], decreasing=TRUE),10))))
#DotPlot(sub, features=sf, group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[1]))  + NoLegend()
p3 <- DotPlot(sub, features="Plp1", group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[1]))

sub <- subset(tepe, subset = ClusterName %in% c(names(head(sort(avg@assays$predictions@scale.data[,clusters[5]], decreasing=TRUE),10))))
#DotPlot(sub, features=sf, group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[1]))  + NoLegend()
p4 <- DotPlot(sub, features="Plp1", group.by = "ClusterName", scale.max=100, col.min = -1.5, col.max = 1.5, cols = c("lightgrey", col[1]))

plot_grid(p1,p2,p3,p3,ncol=4)

tiff("figures/sup.fig8.plp1.tif", units="in", width=18, height=4, res=300)
plot_grid(p1,p2,ncol=5)
dev.off()

tiff("figures/sup.fig8.plp1-2.tif", units="in", width=10, height=4, res=300)
plot_grid(p3,p4,ncol=2)
dev.off()

```

# Supplementary fig.5 - spatial markers

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=4, fig.width=6}

DefaultAssay(hip142079) <- "SCT"
hip142079 <- subset(hip142079, subset = nFeature_ISO>1)
SpatialDimPlot(hip142079, pt.size.factor =1.6, label=TRUE)

markers <- read.delim("./output/hip142079.markers.csv", sep=",", stringsAsFactors = F)
markers <- markers[markers$pct.2<0.5,]
top <- markers %>% group_by(cluster) %>% top_n(10, avg_logFC)
DotPlot(object = hip142079, features = unique(top$gene), scale=TRUE, scale.max = 100, scale.min = 0, col.min = -1, col.max = 2, cols = c("lightgrey", "red")) + RotatedAxis()
SpatialFeaturePlot(object = hip142079, features = c("C1ql2","Myo5b","Slc6a3","Zic1","Fibcd1","Nrep","Mag","Hpcal1","Camk2n1","Myl4","Gfap","Mbp"), pt.size.factor=1.6, alpha = c(1, 1))

tiff("figures/sup.markers.hip.tif", units="in", width=14, height=12, res=300)
SpatialFeaturePlot(object = hip142079, features = c("C1ql2","Myo5b","Slc6a3","Zic1","Fibcd1","Nrep","Mag","Hpcal1","Camk2n1","Myl4","Gfap","Calb2"), pt.size.factor=1.6, alpha = c(1, 1))
dev.off()

tiff("figures/sup.markers.dotplot.hip.tif", units="in", width=10, height=6, res=300)
DotPlot(object = hip142079, features = rev(c("C1ql2","Myo5b","Slc6a3","Zic1","Fibcd1","Nrep","Mag","Hpcal1","Camk2n1","Myl4","Gfap","Mbp")), scale=TRUE, scale.max = 100, scale.min = 0, col.min = -1, col.max = 2, cols = c("lightgrey", "red")) + RotatedAxis()
dev.off()

DefaultAssay(mob164315) <- "SCT"
tiff("figures/mob.tif", units="in", width=6, height=6, res=300)
SpatialDimPlot(mob164315, pt.size.factor =2)
dev.off()

DotPlot(object = mob164315, features = c("Nrgn","Tbx21","Bc1","Eomes","Nnat"), scale=TRUE, scale.max = 100, scale.min = 0, col.min = -1, col.max = 2, cols = c("lightgrey", "red"))

tiff("figures/sup.markers.mob.tif", units="in", width=6, height=6, res=300)
SpatialFeaturePlot(mob164315, features = c("Nrgn","Tbx21","Bc1","Eomes","Nnat"), slot="data", pt.size.factor=1.6, alpha = c(1, 1))
dev.off()

markers <- read.delim("./output/mob164315.markers.csv", sep=",", stringsAsFactors = F)
markers <- markers[markers$pct.2<0.5,]
top <- markers %>% group_by(cluster) %>% top_n(10, avg_logFC)
DotPlot(object = mob164315, features = unique(top$gene), scale=TRUE, scale.max = 100, scale.min = 0, col.min = -1, col.max = 2, cols = c("lightgrey", "red")) + RotatedAxis()

tiff("figures/sup.markers.mob.tif", units="in", width=6, height=6, res=300)
SpatialFeaturePlot(object = mob164315, features = c("Gsn","Trh","Cnot3","Nrgn","Spp1"), pt.size.factor=2, alpha = c(1, 1))
dev.off()

tiff("figures/sup.markers.dotplot.mob.tif", units="in", width=10, height=6, res=300)
DotPlot(object = mob164315, features = rev(c("Gsn","Trh","Cnot3","Nrgn","Spp1")), scale=TRUE, scale.max = 100, scale.min = 0, col.min = -1, col.max = 2, cols = c("lightgrey", "red")) + RotatedAxis()
dev.off()

```

# Editing statistics

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=4, fig.width=12}

length(rownames(hip142079@assays$AtoI@counts))
sum(hip142079[['AtoI_total']], na.rm=TRUE)
sum(hip142079[['AtoI_G']], na.rm=TRUE)
sum(hip142079[['AtoI_A']], na.rm=TRUE)
100*sum(hip142079[['AtoI_G']], na.rm=TRUE)/sum(hip142079[['AtoI_total']], na.rm=TRUE)

```

# Main fig.2x - 86 editing sites heatmap (not retained)

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=4, fig.width=12}

all = read.delim("./nanopore/hip142079.snp100838.RN3.QV5_snpmatrix.full.txt", stringsAsFactors = F)
data = all[,4:ncol(all)]
dim(data)
colnames(data) <- paste(colnames(data),"-1", sep="")
rownames(data) = paste(all$geneId, all$transcriptId, sep=":")
vect <- colnames(hip142079@assays$Spatial)
#idx <- grep("intergenic", rownames(data), invert = TRUE)
#data <- data[idx,vect]
#idx <- grep("Rik", rownames(data), invert = TRUE)
data <- data[,vect]
idx1 <- grep("\\.\\.A", rownames(data))
idx2 <- grep("\\.\\.G", rownames(data))
data <- data[c(idx1,idx2),]
dim(data)

x <- unique(sapply(strsplit(rownames(data), "\\.\\."), `[`, 1))
countA <- data[grep("\\.\\.A",rownames(data)),]
colnames(countA) <- colnames(data)
rownames(countA) <- x
countG <- data[grep("\\.\\.G",rownames(data)),]
colnames(countG) <- colnames(data)
rownames(countG) <- x

countG <- countG[rowSums(countG)>49,]
countA <- countA[rownames(countG),]
dim(countA)
dim(countG)

hip142079[["AtoIhigh"]] <- CreateAssayObject(counts = countA)
hip142079 <- SetAssayData(hip142079, slot = "data", as.matrix(countG), assay = "AtoIhigh")

agg.A <- data.frame()
for(i in 1:86){
  t <- aggregate(hip142079@assays$AtoIhigh@counts[i,], list(hip142079@meta.data$ClusterName), sum)
  agg.A <- rbind(agg.A,t$x)
  print(i)
}
colnames(agg.A) <- unique(t$Group.1)
rownames(agg.A) <- rownames(hip142079@assays$AtoIhigh@counts)

agg.G <- data.frame()
for(i in 1:86){
  t <- aggregate(hip142079@assays$AtoIhigh@data[i,], list(hip142079@meta.data$ClusterName), sum)
  agg.G <- rbind(agg.G,t$x)
  print(i)
}
colnames(agg.G) <- unique(t$Group.1)
rownames(agg.G) <- rownames(hip142079@assays$AtoIhigh@data)

ratio <- agg.G/(agg.A+agg.G)
ratio[is.na(ratio)] <- 0
dim(ratio)

pheatmap(as.matrix(ratio), cluster_rows=TRUE, cluster_cols=FALSE, fontsize = 8, clustering_method = "complete", clustering_distance_rows = "manhattan")

avg[["AtoIhigh"]] <- CreateAssayObject(counts = ratio)
VlnPlot(avg,features="Calm1:12:100207186", slot="data",assay="AtoIhigh")
VlnPlot(avg,features="Scn1b:7:31116973", slot="data",assay="AtoIhigh")

df <- as.data.frame(avg@assays$AtoIhigh@counts[rownames(ratio[grep("^Gria2", rownames(ratio)),]),])
df$id <- rownames(df)
df.molten <- melt(df, id.vars="id")
ggplot(df.molten, aes(x=variable, y=value, fill=id)) +
    geom_bar(stat='identity', position='dodge') +
    theme_minimal()

df <- as.data.frame(avg@assays$AtoIhigh@counts[rownames(ratio[grep("^Blcap", rownames(ratio)),]),])
df$id <- rownames(df)
df.molten <- melt(df, id.vars="id")
ggplot(df.molten, aes(x=variable, y=value, fill=id)) +
    geom_bar(stat='identity', position='dodge') +
    theme_minimal()

df <- as.data.frame(avg@assays$AtoIhigh@counts[rownames(ratio[grep("^Grik5", rownames(ratio)),]),])
df$id <- rownames(df)
df.molten <- melt(df, id.vars="id")
ggplot(df.molten, aes(x=variable, y=value, fill=id)) +
    geom_bar(stat='identity', position='dodge') +
    theme_minimal()

df <- as.data.frame(avg@assays$AtoIhigh@counts[rownames(ratio[grep("^Rpa1", rownames(ratio)),]),])
df$id <- rownames(df)
df.molten <- melt(df, id.vars="id")
ggplot(df.molten, aes(x=variable, y=value, fill=id)) +
    geom_bar(stat='identity', position='dodge') +
    theme_minimal()

#write.table(agg.A, file="hip142079.A.csv", sep=",")
#write.table(agg.G, file="hip142079.G.csv", sep=",")
#write.table(ratio, file="hip142079.R.csv", sep=",")

```

# Main fig.2d - Calm1 editing

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=4, fig.width=12}

gene <- "Calm1"
x <- unique(rownames(hip142079@assays$AtoI@counts)[grep(paste0("^",gene,":"), rownames(hip142079@assays$AtoI@counts))])
DefaultAssay(hip142079) <- "AtoI"
p1 <- SpatialFeaturePlot(hip142079, features = x, slot="counts", pt.size.factor=1.6, alpha = c(1, 1)) + labs(title = "Non edited molecules")
p2 <- SpatialFeaturePlot(hip142079, features = x, slot="data", pt.size.factor=1.6, alpha = c(1, 1), max.cutoff=c(7)) + labs(title = "Edited molecules")
p3 <- SpatialFeaturePlot(hip142079, features = x, slot="scale.data", pt.size.factor=1.6, alpha = c(1, 1), max.cutoff=c(0.5)) + labs(title = "Editing ratio")

hip142079[['editing_ratio']] <- hip142079@assays$AtoI@scale.data[x,]
p4 <- ggplot(data=hip142079@meta.data, aes(x=ClusterName,y=editing_ratio, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  #geom_text(aes(label=geneCount), vjust=1.5, color="black", position = position_dodge(0.9), size=3)+
  scale_y_continuous(limits=c(0,0.4), breaks=seq(0,0.4,0.1), expand = c(0, 0)) +
  coord_flip() + labs(title = "\n\nEditing ratio per region\n")

plot_grid(p1,p2,p3,p4, ncol=4)

```

# Scn1b editing

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=4, fig.width=12}

gene <- "Scn1b"
x <- unique(rownames(hip142079@assays$AtoI@counts)[grep(paste0("^",gene,":"), rownames(hip142079@assays$AtoI@counts))])
DefaultAssay(hip142079) <- "AtoI"
p1 <- SpatialFeaturePlot(hip142079, features = x, slot="counts", pt.size.factor=2, alpha = c(1, 1)) + labs(title = "Non edited molecules")
p2 <- SpatialFeaturePlot(hip142079, features = x, slot="data", pt.size.factor=2, alpha = c(1, 1)) + labs(title = "Edited molecules")
p3 <- SpatialFeaturePlot(hip142079, features = x, slot="scale.data", pt.size.factor=2, alpha = c(1, 1)) + labs(title = "Editing ratio")

RidgePlot(hip142079,x,slot="scale.data")
VlnPlot(hip142079,x,slot="scale.data")

hip142079[['editing_ratio']] <- hip142079@assays$AtoI@scale.data[x,]
p4 <- ggplot(data=hip142079@meta.data, aes(x=ClusterName,y=editing_ratio, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  #geom_text(aes(label=geneCount), vjust=1.5, color="black", position = position_dodge(0.9), size=3)+
  #scale_y_continuous(limits=c(0,0.4), breaks=seq(0,0.4,0.1), expand = c(0, 0)) +
  coord_flip() + labs(title = "\n\nEditing ratio per region\n")

plot_grid(p1,p2,p3,p4, ncol=4)

```

# Gnas editing

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=4, fig.width=12}

gene <- "Gnas"
x <- unique(rownames(hip142079@assays$AtoI@counts)[grep(paste0("^",gene,":"), rownames(hip142079@assays$AtoI@counts))])
DefaultAssay(hip142079) <- "AtoI"
p1 <- SpatialFeaturePlot(hip142079, features = x, slot="counts", pt.size.factor=2, alpha = c(1, 1)) + labs(title = "Non edited molecules")
p2 <- SpatialFeaturePlot(hip142079, features = x, slot="data", pt.size.factor=2, alpha = c(1, 1)) + labs(title = "Edited molecules")
p3 <- SpatialFeaturePlot(hip142079, features = x, slot="scale.data", pt.size.factor=2, alpha = c(1, 1)) + labs(title = "Editing ratio")

hip142079[['editing_ratio']] <- hip142079@assays$AtoI@scale.data[x,]
p4 <- ggplot(data=hip142079@meta.data, aes(x=ClusterName,y=editing_ratio, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  #geom_text(aes(label=geneCount), vjust=1.5, color="black", position = position_dodge(0.9), size=3)+
  scale_y_continuous(limits=c(0,0.4), breaks=seq(0,0.4,0.1), expand = c(0, 0)) +
  coord_flip() + labs(title = "\n\nEditing ratio per region\n")

plot_grid(p1,p2,p3,p4, ncol=4)

```

# Distinct number of editing sites detected as edited by at least one UMI (I)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

p1 <- SpatialFeaturePlot(hip142079, features = "AtoI_detected", pt.size.factor=2, alpha = c(1, 1))
p2 <- VlnPlot(hip142079,"AtoI_detected")
p3 <- RidgePlot(hip142079, features = "AtoI_detected")

plot_grid(p1,p2,p3,ncol=3)

p1 <- ggplot(data=hip142079@meta.data, aes(x=ClusterName, y=AtoI_A, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  #geom_text(aes(label=geneCount), vjust=1.5, color="black", position = position_dodge(0.9), size=3)+
  #scale_y_continuous(limits=c(0,40), breaks=seq(0,40,10), expand = c(0, 0)) +
  labs(title = "AI1886_A")

p2 <- ggplot(data=hip142079@meta.data, aes(x=ClusterName, y=AtoI_G, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  #geom_text(aes(label=geneCount), vjust=1.5, color="black", position = position_dodge(0.9), size=3)+
  #scale_y_continuous(limits=c(0,40), breaks=seq(0,40,10), expand = c(0, 0)) +
  labs(title = "AI1886_G")

p3 <- ggplot(data=hip142079@meta.data, aes(x=ClusterName, y=AtoI_ratio, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  #geom_text(aes(label=geneCount), vjust=1.5, color="black", position = position_dodge(0.9), size=3)+
  #scale_y_continuous(limits=c(0,40), breaks=seq(0,40,10), expand = c(0, 0)) +
  labs(title = "AI1886_ratio")

p4 <- ggplot(data=hip142079@meta.data, aes(x=ClusterName, y=AtoI_detected, group=ClusterName))+
  geom_boxplot(outlier.shape = NA, fill=col, color="black") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  #geom_text(aes(label=geneCount), vjust=1.5, color="black", position = position_dodge(0.9), size=3)+
  #scale_y_continuous(limits=c(0,40), breaks=seq(0,40,10), expand = c(0, 0)) +
  labs(title = "AI1886_detected")

plot_grid(p1,p2,p3,p4,ncol=4)

```

# export tables

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=4, fig.width=6}

hip142079 <- readRDS("./output/hip142079.nanopore.rds")
df <- data.frame()
df <- data.frame(Matrix::rowSums(hip142079@assays$AtoI@counts), Matrix::rowSums(hip142079@assays$AtoI@data))
colnames(df) <- c("A", "G")
df$total <- df$A+df$G
df$ratio <- df$G/df$total
write.table(df, file="hip142079.AtoI.csv", sep=",")

hip142078 <- readRDS("./output/hip142078.nanopore.rds")
df <- data.frame()
df <- data.frame(Matrix::rowSums(hip142078@assays$AtoI@counts), Matrix::rowSums(hip142078@assays$AtoI@data))
colnames(df) <- c("A", "G")
df$total <- df$A+df$G
df$ratio <- df$G/df$total
write.table(df, file="hip142078.AtoI.csv", sep=",")

mob164315 <- readRDS("./output/mob164315.nanopore.rds")
df <- data.frame()
df <- data.frame(Matrix::rowSums(mob164315@assays$AtoI@counts), Matrix::rowSums(mob164315@assays$AtoI@data))
colnames(df) <- c("A", "G")
df$total <- df$A+df$G
df$ratio <- df$G/df$total
write.table(df, file="mob164315.AtoI.csv", sep=",")

```

# Supplementary fig.13 - Nanopore/Illumina agreement

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

library(viridis)

agreement = read.delim("ressources/agreement.txt", sep="\t", stringsAsFactors = F)
ggplot(data=agreement, aes(x=RN, y=QV, size=number, fill=agree)) +
    geom_point(alpha=0.5, shape=21, color="black") + 
    scale_size(range = c(.1, 36), name="total") +
    scale_fill_viridis(guide=FALSE, option="A") +
    theme_minimal() +
    theme(legend.position="bottom") +
    ylab("Nanopore base QV") +
    xlab("Molecule read number (RN)") +
    theme(legend.position = "none")

tiff("figures/RNagree.tif", units="in", width=6, height=8, res=300)
agreement = read.delim("ressources/RNAgree.txt", sep="\t", stringsAsFactors = F)
ggplot(data=agreement, aes(x=RN, y=Agreement)) +
    geom_line()+
    geom_point()
dev.off()

tiff("figures/QVagree.tif", units="in", width=6, height=8, res=300)
agreement = read.delim("ressources/QVAgree.txt", sep="\t", stringsAsFactors = F)
ggplot(data=agreement, aes(x=QV, y=Agreement)) +
    geom_line()+
  theme_minimal() +
    geom_point()
dev.off()


```

# Supplementary fig.9 - hip78/hip79 correlation

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

breaksList = seq(0.6, 1, by = 0.01)

hip142078 <- subset(hip142078, ident=c("Midbrain", "Hippocampus", "Isocortex-1", "Isocortex-2", "Olfactory area", "Fiber tracts", "Isocortex-3", "Hammons horn", "Thalamus", "Hypothalamus", "Ca3", "Dg"))

p1 <- SpatialDimPlot(hip142078)
p2 <- SpatialDimPlot(hip142079)
plot_grid(p1,p2)

# Spatial correlation
avg79 <- AverageExpression(object = hip142079, return.seurat = TRUE, assay="SCT")
avg78 <- AverageExpression(object = hip142078, return.seurat = TRUE, assay="SCT")

regions <- rownames(avg79@meta.data)
genes <- intersect(rownames(avg78@assays$SCT@data), rownames(avg79@assays$SCT@data))
length(genes)

df <- data.frame()
for (i in (1:length(regions))){
  
  df2<- data.frame()
  for (j in (1:length(regions))){
    t <- data.frame(avg78@assays$SCT@data[genes,regions[i]], avg79@assays$SCT@data[genes,regions[j]])
    
    colnames(t) <- c("avg78","avg79")
    c <- cor.test(t$avg78, t$avg79)$estimate
    
    df[i,j] <-c
  }
}
colnames(df) <- c("Midbrain", "Hippocampus", "Isocortex-1", "Isocortex-2", "Olfactory area", "Fiber tracts", "Isocortex-3", "Hammons horn", "Thalamus", "Hypothalamus", "Ca3", "Dg")
rownames(df) <- c("Midbrain", "Hippocampus", "Isocortex-1", "Isocortex-2", "Olfactory area", "Fiber tracts", "Isocortex-3", "Hammons horn", "Thalamus", "Hypothalamus", "Ca3", "Dg")

tiff("figures/Cor.region.spatial.tif", units="in", width=4, height=3, res=300)
pheatmap(df, cluster_rows=FALSE, cluster_cols=FALSE, fontsize = 8, clustering_method = "complete", clustering_distance_rows = "manhattan", color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)), breaks = breaksList)
dev.off()

# Nanopore ISOG correlation
avg79 <- AverageExpression(object = hip142079, return.seurat = TRUE, assay="ISOG")
avg78 <- AverageExpression(object = hip142078, return.seurat = TRUE, assay="ISOG")

genes <- intersect(rownames(avg78@assays$ISOG@data), rownames(avg79@assays$ISOG@data))
length(genes)

df <- data.frame()
for (i in (1:length(regions))){
  
  df2<- data.frame()
  for (j in (1:length(regions))){
    t <- data.frame(avg78@assays$ISOG@data[genes,regions[i]], avg79@assays$ISOG@data[genes,regions[j]])
    
    colnames(t) <- c("avg78","avg79")
    c <- cor.test(t$avg78, t$avg79)$estimate
    
    df[i,j] <-c
  }
}
colnames(df) <- c("Midbrain", "Hippocampus", "Isocortex-1", "Isocortex-2", "Olfactory area", "Fiber tracts", "Isocortex-3", "Hammons horn", "Thalamus", "Hypothalamus", "Ca3", "Dg")
rownames(df) <- c("Midbrain", "Hippocampus", "Isocortex-1", "Isocortex-2", "Olfactory area", "Fiber tracts", "Isocortex-3", "Hammons horn", "Thalamus", "Hypothalamus", "Ca3", "Dg")

tiff("figures/Cor.region.ISOG.tif", units="in", width=4, height=3, res=300)
pheatmap(df, cluster_rows=FALSE, cluster_cols=FALSE, fontsize = 8, clustering_method = "complete", clustering_distance_rows = "manhattan", color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)), breaks = breaksList)
dev.off()

# Nanopore ISO correlation
avg79 <- AverageExpression(object = hip142079, return.seurat = TRUE, assay="ISO")
avg78 <- AverageExpression(object = hip142078, return.seurat = TRUE, assay="ISO")

genes <- intersect(rownames(avg78@assays$ISO@data), rownames(avg79@assays$ISO@data))
length(genes)

df <- data.frame()
for (i in (1:length(regions))){
  
  df2<- data.frame()
  for (j in (1:length(regions))){
    t <- data.frame(avg78@assays$ISO@data[genes,regions[i]], avg79@assays$ISO@data[genes,regions[j]])
    
    colnames(t) <- c("avg78","avg79")
    c <- cor.test(t$avg78, t$avg79)$estimate
    
    df[i,j] <-c
  }
}
colnames(df) <- c("Midbrain", "Hippocampus", "Isocortex-1", "Isocortex-2", "Olfactory area", "Fiber tracts", "Isocortex-3", "Hammons horn", "Thalamus", "Hypothalamus", "Ca3", "Dg")
rownames(df) <- c("Midbrain", "Hippocampus", "Isocortex-1", "Isocortex-2", "Olfactory area", "Fiber tracts", "Isocortex-3", "Hammons horn", "Thalamus", "Hypothalamus", "Ca3", "Dg")

tiff("figures/Cor.region.ISO.tif", units="in", width=4, height=3, res=300)
pheatmap(df, cluster_rows=FALSE, cluster_cols=FALSE, fontsize = 8, clustering_method = "complete", clustering_distance_rows = "manhattan", color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)), breaks = breaksList)
dev.off()

```

# Session Info

```{r sessinf}
sessionInfo()
```

