---
title: "Visium Spatial Tx - 10.correlation"
author: '[Kevin Lebrigand](mailto:lebrigand@ipmc.cnrs.fr) CNRS Senior Scientist - Computational Biology'
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  html_notebook:
    code_folding: none
    theme: journal
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

<style type="text/css">

body, td {
   font-size: 15px;
}
code.r{
  font-size: 15px;
}
pre {
  font-size: 15px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  cache.lazy = FALSE,
  tidy = TRUE
)
```

# Loading data

```{r message=FALSE, warning=FALSE, echo=FALSE, eval=T, fig.height=4, fig.width=6}

setwd("/data/10x_data/10x_visium/")
library("reshape2")
library(gameofthrones)
source("00.import.R")

```

# mob164315

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

visium <- readRDS("./output/mob164315.nanopore.rds")
col=ggplotColours(n = length(unique(visium@active.ident)))

median(visium@meta.data$nCount_Spatial)
median(visium@meta.data$nCount_ISOG)
median(visium@meta.data$nCount_ISO)

median(visium@meta.data$nFeature_Spatial)
median(visium@meta.data$nFeature_ISOG)
median(visium@meta.data$nFeature_ISO)

```

## Percentage Illumina UMIs / Features seen in Nanopore data

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

visium@meta.data$nCount_profiled <- visium@meta.data$nCount_ISOG*100/visium@meta.data$nCount_Spatial
mean(visium@meta.data$nCount_profiled)
visium@meta.data$nFeature_profiled <- visium@meta.data$nFeature_ISOG*100/visium@meta.data$nFeature_Spatial
mean(visium@meta.data$nFeature_profiled)

```

## Correlation illumina/Nanopore (r²)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

#pdf("Correlation.plots.pdf", width=10, height=8, useDingbats=FALSE)

p1 <- ggplot(data=visium@meta.data, aes(x=nFeature_Spatial,y=nFeature_ISOG)) +  geom_point(shape = 21, colour = "black", fill = col[visium@active.ident]) +
      theme_bw() + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
      ggtitle("Features per cell") +
      #geom_abline(slope=1, intercept=-15, linetype=3) +
      labs(x="Illumina",y="Nanopore")

x <- cor.test(visium@meta.data$nFeature_Spatial, visium@meta.data$nFeature_ISOG, method="pearson")$estimate
x*x

p2 <- ggplot(data=visium@meta.data, aes(x=nCount_Spatial,y=nCount_ISOG)) +  geom_point(shape = 21, colour = "black", fill = col[visium@active.ident]) +
      theme_bw() + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
      ggtitle("UMIs per cell") +
      #geom_abline(slope=1, intercept=-15, linetype=3) +
      labs(x="Illumina",y="Nanopore")

x <- cor.test(visium@meta.data$nCount_Spatial, visium@meta.data$nCount_ISOG, method="pearson")$estimate
x*x

tiff("figures/scatterplot.mob.tif", units="in", width=4, height=8, res=300)
plot_grid(p1, p2, ncol=1)
dev.off()

plot1 <- ggplot(data=visium@meta.data, aes(nFeature_Spatial)) + 
                geom_histogram(breaks=seq(0, 8000, by = 100),col="black",aes(fill=nGene),fill="lightsteelblue3") + 
                scale_fill_gradient("Count", low = "green", high = "red") +
                labs(title="Illumina nFeature_Spatial") + 
                labs(x="nFeatures", y="Cells")

plot2 <- ggplot(data=visium@meta.data, aes(nFeature_ISOG)) + 
                geom_histogram(breaks=seq(0, 8000, by = 100),col="black",aes(fill=nGene),fill="lightsteelblue3") + 
                scale_fill_gradient("Count", low = "green", high = "red") +
                labs(title="Nanopore nFeature_ISOG") + 
                labs(x="nFeatures", y="Cells")

plot_grid(plot1, plot2)

#dev.off()

```

## Correlation Illumina / Nanopore (r)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

common.cells <- rownames(visium@meta.data)
common.genes <- intersect(rownames(visium@assays$Spatial@data), rownames(visium@assays$ISOG@data))
common.genes.nomt <- common.genes[grep("mt-", common.genes, invert = TRUE)]
common.genes.nomtrp <- common.genes.nomt[grep("Rp[ls]", common.genes.nomt, invert = TRUE)]
length(common.genes.nomtrp)

df <- data.frame()
for (i in (1:length(common.cells))){
  c <- intersect(names(which(visium@assays$Spatial@data[common.genes.nomtrp,common.cells[i]]>0)), names(which(visium@assays$ISOG@data[common.genes.nomtrp,common.cells[i]] > 0)))
  #print(paste(i,common.cells[i],length(c),sep=","))
  t <- data.frame(visium@assays$Spatial@data[c,common.cells[i]], visium@assays$ISOG@data[c,common.cells[i]])
  colnames(t) <- c("illu","nano")
  
  if(length(c)>100){
    df <- rbind(df, data.frame("Illu/ONT same cell",cor.test(t$illu, t$nano)$estimate))
  }
}
colnames(df) <- c("type","cor")
rownames(df) <- c()
median(df$cor, na.rm=TRUE)

dfmob.1 <- df

df2 <- data.frame()
for (i in (1:length(common.cells))){
  #print (i)
  random <- sample(1:length(common.cells), 2, replace=T)
  
  for (j in (1:length(random))){
    if(visium@meta.data$ClusterName[i] != visium@meta.data$ClusterName[random[j]]){
      c <- intersect(names(which(visium@assays$Spatial@data[common.genes.nomtrp,common.cells[random[j]]]>0)), names(which(visium@assays$ISOG@data[common.genes.nomtrp,common.cells[i]] > 0)))
      t <- data.frame(visium@assays$Spatial@data[c,common.cells[random[j]]], visium@assays$ISOG@data[c,common.cells[i]])
      colnames(t) <- c("illu","nano")
      #print(paste(i,common.cells[i],length(c),sep=","))
      if(length(c)>100){
        df2 <- rbind(df2, data.frame("Illu/ONT different cell",cor.test(t$illu, t$nano)$estimate))
      }
    }
  }
}
colnames(df2) <- c("type","cor")
rownames(df2) <- c()
median(df2$cor, na.rm=TRUE)

dfmob.2 <- df2

```
Nanopore/Illumina same cell
```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}
median(df$cor)
```
Nanopore/Illumina different cell
```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}
median(df2$cor)
```

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=6, fig.width=12}

whole <- rbind(df,df2)
#pdf("figures/fig.2c.pdf", width=10, height=8, useDingbats=FALSE)
ggplot(whole, aes(x=type, y=cor, fill=type)) +
  geom_boxplot(outlier.shape = NA)+
  ggtitle("Transcriptomics correlation") +
  theme_classic()
#dev.off()

```

# hip142078

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

visium <- readRDS("./output/hip142078.nanopore.rds")
col=ggplotColours(n = length(unique(visium@active.ident)))

median(visium@meta.data$nCount_Spatial)
median(visium@meta.data$nCount_ISOG)
median(visium@meta.data$nCount_ISO)

median(visium@meta.data$nFeature_Spatial)
median(visium@meta.data$nFeature_ISOG)
median(visium@meta.data$nFeature_ISO)

```

## Percentage Illumina UMIs / Features seen in Nanopore data

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

visium@meta.data$nCount_profiled <- visium@meta.data$nCount_ISOG*100/visium@meta.data$nCount_Spatial
mean(visium@meta.data$nCount_profiled)
visium@meta.data$nFeature_profiled <- visium@meta.data$nFeature_ISOG*100/visium@meta.data$nFeature_Spatial
mean(visium@meta.data$nFeature_profiled)

```

## Correlation illumina/Nanopore (r²)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

#pdf("Correlation.plots.pdf", width=10, height=8, useDingbats=FALSE)

p1 <- ggplot(data=visium@meta.data, aes(x=nFeature_Spatial,y=nFeature_ISOG)) +  geom_point(shape = 21, colour = "black", fill = col[visium@active.ident]) +
      theme_bw() + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
      ggtitle("Features per cell") +
      #geom_abline(slope=1, intercept=-15, linetype=3) +
      labs(x="Illumina",y="Nanopore")

x <- cor.test(visium@meta.data$nFeature_Spatial, visium@meta.data$nFeature_ISOG, method="pearson")$estimate
x*x

p2 <- ggplot(data=visium@meta.data, aes(x=nCount_Spatial,y=nCount_ISOG)) +  geom_point(shape = 21, colour = "black", fill = col[visium@active.ident]) +
      theme_bw() + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
      ggtitle("UMIs per cell") +
      #geom_abline(slope=1, intercept=-15, linetype=3) +
      labs(x="Illumina",y="Nanopore")

x <- cor.test(visium@meta.data$nCount_Spatial, visium@meta.data$nCount_ISOG, method="pearson")$estimate
x*x

tiff("figures/scatterplot.hip78.tif", units="in", width=4, height=8, res=300)
plot_grid(p1, p2, ncol=1)
dev.off()


plot1 <- ggplot(data=visium@meta.data, aes(nFeature_Spatial)) + 
                geom_histogram(breaks=seq(0, 8000, by = 100),col="black",aes(fill=nGene),fill="lightsteelblue3") + 
                scale_fill_gradient("Count", low = "green", high = "red") +
                labs(title="Illumina nFeature_Spatial") + 
                labs(x="nFeatures", y="Cells")

plot2 <- ggplot(data=visium@meta.data, aes(nFeature_ISOG)) + 
                geom_histogram(breaks=seq(0, 8000, by = 100),col="black",aes(fill=nGene),fill="lightsteelblue3") + 
                scale_fill_gradient("Count", low = "green", high = "red") +
                labs(title="Nanopore nFeature_ISOG") + 
                labs(x="nFeatures", y="Cells")

plot_grid(plot1, plot2)

#dev.off()

```

## Correlation Illumina / Nanopore (r)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

common.cells <- rownames(visium@meta.data)
common.genes <- intersect(rownames(visium@assays$Spatial@data), rownames(visium@assays$ISOG@data))
common.genes.nomt <- common.genes[grep("mt-", common.genes, invert = TRUE)]
common.genes.nomtrp <- common.genes.nomt[grep("Rp[ls]", common.genes.nomt, invert = TRUE)]
length(common.genes.nomtrp)

df <- data.frame()
for (i in (1:length(common.cells))){
  c <- intersect(names(which(visium@assays$Spatial@data[common.genes.nomtrp,common.cells[i]]>0)), names(which(visium@assays$ISOG@data[common.genes.nomtrp,common.cells[i]] > 0)))
  #print(paste(i,common.cells[i],length(c),sep=","))
  t <- data.frame(visium@assays$Spatial@data[c,common.cells[i]], visium@assays$ISOG@data[c,common.cells[i]])
  colnames(t) <- c("illu","nano")
  
  if(length(c)>100){
    df <- rbind(df, data.frame("Illu/ONT same cell",cor.test(t$illu, t$nano)$estimate))
  }
}
colnames(df) <- c("type","cor")
rownames(df) <- c()
median(df$cor, na.rm=TRUE)

dfhip8.1 <- df

df2 <- data.frame()
for (i in (1:length(common.cells))){
  #print (i)
  random <- sample(1:length(common.cells), 2, replace=T)
  
  for (j in (1:length(random))){
    if(visium@meta.data$ClusterName[i] != visium@meta.data$ClusterName[random[j]]){
      c <- intersect(names(which(visium@assays$Spatial@data[common.genes.nomtrp,common.cells[random[j]]]>0)), names(which(visium@assays$ISOG@data[common.genes.nomtrp,common.cells[i]] > 0)))
      t <- data.frame(visium@assays$Spatial@data[c,common.cells[random[j]]], visium@assays$ISOG@data[c,common.cells[i]])
      colnames(t) <- c("illu","nano")
      #print(paste(i,common.cells[i],length(c),sep=","))
      if(length(c)>100){
        df2 <- rbind(df2, data.frame("Illu/ONT different cell",cor.test(t$illu, t$nano)$estimate))
      }
    }
  }
}
colnames(df2) <- c("type","cor")
rownames(df2) <- c()
median(df2$cor, na.rm=TRUE)

dfhip8.2 <- df2

```
Nanopore/Illumina same cell
```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}
median(df$cor)
```
Nanopore/Illumina different cell
```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}
median(df2$cor)
```

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=6, fig.width=12}

whole <- rbind(df,df2)
#pdf("figures/fig.2c.pdf", width=10, height=8, useDingbats=FALSE)
ggplot(whole, aes(x=type, y=cor, fill=type)) +
  geom_boxplot(outlier.shape = NA)+
  ggtitle("Transcriptomics correlation") +
  theme_classic()
#dev.off()

```

# hip142079

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

visium <- readRDS("./output/hip142079.nanopore.rds")
col=ggplotColours(n = length(unique(visium@active.ident)))

median(visium@meta.data$nCount_Spatial)
median(visium@meta.data$nCount_ISOG)
median(visium@meta.data$nCount_ISO)

median(visium@meta.data$nFeature_Spatial)
median(visium@meta.data$nFeature_ISOG)
median(visium@meta.data$nFeature_ISO)

```

## Percentage Illumina UMIs / Features seen in Nanopore data

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

visium@meta.data$nCount_profiled <- visium@meta.data$nCount_ISOG*100/visium@meta.data$nCount_Spatial
mean(visium@meta.data$nCount_profiled)
visium@meta.data$nFeature_profiled <- visium@meta.data$nFeature_ISOG*100/visium@meta.data$nFeature_Spatial
mean(visium@meta.data$nFeature_profiled)

```

## Correlation illumina/Nanopore (r²)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

#pdf("Correlation.plots.pdf", width=10, height=8, useDingbats=FALSE)

p1 <- ggplot(data=visium@meta.data, aes(x=nFeature_Spatial,y=nFeature_ISOG)) +  geom_point(shape = 21, colour = "black", fill = col[visium@active.ident]) +
      theme_bw() + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
      ggtitle("Features per cell") +
      #geom_abline(slope=1, intercept=-15, linetype=3) +
      labs(x="Illumina",y="Nanopore")

x <- cor.test(visium@meta.data$nFeature_Spatial, visium@meta.data$nFeature_ISOG, method="pearson")$estimate
x*x

p2 <- ggplot(data=visium@meta.data, aes(x=nCount_Spatial,y=nCount_ISOG)) +  geom_point(shape = 21, colour = "black", fill = col[visium@active.ident]) +
      theme_bw() + theme(panel.grid.minor=element_blank(), panel.grid.major=element_blank()) +
      ggtitle("UMIs per cell") +
      #geom_abline(slope=1, intercept=-15, linetype=3) +
      labs(x="Illumina",y="Nanopore")

x <- cor.test(visium@meta.data$nCount_Spatial, visium@meta.data$nCount_ISOG, method="pearson")$estimate
x*x

tiff("figures/scatterplot.hip79.tif", units="in", width=4, height=8, res=300)
plot_grid(p1, p2, ncol=1)
dev.off()


plot1 <- ggplot(data=visium@meta.data, aes(nFeature_Spatial)) + 
                geom_histogram(breaks=seq(0, 8000, by = 100),col="black",aes(fill=nGene),fill="lightsteelblue3") + 
                scale_fill_gradient("Count", low = "green", high = "red") +
                labs(title="Illumina nFeature_Spatial") + 
                labs(x="nFeatures", y="Cells")

plot2 <- ggplot(data=visium@meta.data, aes(nFeature_ISOG)) + 
                geom_histogram(breaks=seq(0, 8000, by = 100),col="black",aes(fill=nGene),fill="lightsteelblue3") + 
                scale_fill_gradient("Count", low = "green", high = "red") +
                labs(title="Nanopore nFeature_ISOG") + 
                labs(x="nFeatures", y="Cells")

plot_grid(plot1, plot2)

#dev.off()

```

## Correlation Illumina / Nanopore (r)

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}

common.cells <- rownames(visium@meta.data)
common.genes <- intersect(rownames(visium@assays$Spatial@data), rownames(visium@assays$ISOG@data))
common.genes.nomt <- common.genes[grep("mt-", common.genes, invert = TRUE)]
common.genes.nomtrp <- common.genes.nomt[grep("Rp[ls]", common.genes.nomt, invert = TRUE)]
length(common.genes.nomtrp)

df <- data.frame()
for (i in (1:length(common.cells))){
  c <- intersect(names(which(visium@assays$Spatial@data[common.genes.nomtrp,common.cells[i]]>0)), names(which(visium@assays$ISOG@data[common.genes.nomtrp,common.cells[i]] > 0)))
  #print(paste(i,common.cells[i],length(c),sep=","))
  t <- data.frame(visium@assays$Spatial@data[c,common.cells[i]], visium@assays$ISOG@data[c,common.cells[i]])
  colnames(t) <- c("illu","nano")
  
  if(length(c)>100){
    df <- rbind(df, data.frame("Illu/ONT same cell",cor.test(t$illu, t$nano)$estimate))
  }
}
colnames(df) <- c("type","cor")
rownames(df) <- c()
median(df$cor, na.rm=TRUE)

dfhip9.1 <- df

df2 <- data.frame()
for (i in (1:length(common.cells))){
  #print (i)
  random <- sample(1:length(common.cells), 2, replace=T)
  
  for (j in (1:length(random))){
    if(visium@meta.data$ClusterName[i] != visium@meta.data$ClusterName[random[j]]){
      c <- intersect(names(which(visium@assays$Spatial@data[common.genes.nomtrp,common.cells[random[j]]]>0)), names(which(visium@assays$ISOG@data[common.genes.nomtrp,common.cells[i]] > 0)))
      t <- data.frame(visium@assays$Spatial@data[c,common.cells[random[j]]], visium@assays$ISOG@data[c,common.cells[i]])
      colnames(t) <- c("illu","nano")
      #print(paste(i,common.cells[i],length(c),sep=","))
      if(length(c)>100){
        df2 <- rbind(df2, data.frame("Illu/ONT different cell",cor.test(t$illu, t$nano)$estimate))
      }
    }
  }
}
colnames(df2) <- c("type","cor")
rownames(df2) <- c()
median(df2$cor, na.rm=TRUE)

dfhip9.2 <- df2

```
Nanopore/Illumina same cell
```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}
median(df$cor)
```
Nanopore/Illumina different cell
```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=4, fig.width=6}
median(df2$cor)
```

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=6, fig.width=12}

whole <- rbind(df,df2)
#pdf("figures/fig.2c.pdf", width=10, height=8, useDingbats=FALSE)
ggplot(whole, aes(x=type, y=cor, fill=type)) +
  geom_boxplot(outlier.shape = NA)+
  ggtitle("Transcriptomics correlation") +
  theme_classic()
#dev.off()

```

# general boxplot

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=T, fig.height=6, fig.width=12}

dfhip9.1$type <- "dfhip9.same"
dfhip9.2$type <- "dfhip9.diff"
dfhip8.1$type <- "dfhip8.same"
dfhip8.2$type <- "dfhip8.diff"
dfmob.1$type <- "dfmob.same"
dfmob.2$type <- "dfmob.diff"

whole <- rbind(dfhip9.1,dfhip8.1,dfmob.1,dfhip9.2,dfhip8.2,dfmob.2)
#pdf("figures/fig.2c.pdf", width=10, height=8, useDingbats=FALSE)
ggplot(whole, aes(x=type, y=cor, fill=type)) +
  geom_boxplot(outlier.shape = NA)+
  ggtitle("Transcriptomics correlation") +
  theme_classic()

```

# Session Info

```{r sessinf}
sessionInfo()
```
